---
title: "Exploratory Analysis of Bakery Data"
author: "Andrew Whang"
date: "October 26, 2018"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Import data and Packages
Load packages, lubridate, dplyr, ggplot2, bindrcpp.
Data was found from Kaggle at <https://www.kaggle.com/xvivancos/transactions-from-a-bakery>
```{r message = FALSE, warning = FALSE}
library(lubridate)
library(dplyr)
library(ggplot2)
library(bindrcpp)
df <- read.csv("BreadBasket_DMS.csv")
```

### Date and Time Conversions
Convert Date and Time to proper formats
Also add columns for day of the week and the month of the transaction

```{r}
df$Date <- as.Date(df$Date)
df$Time <- hms(df$Time)
df$weekday <- wday(df$Date, label = TRUE)
df$month <- month(df$Date, label = TRUE)
```
### Note on filters
We also remove items labeled as "NONE" and select the top 15 most sold items
```{r message = FALSE, echo = FALSE}
top_item <- df %>% 
  select(Item) %>% 
  filter(Item != "NONE") %>% 
  group_by(Item) %>% 
  tally() %>% 
  top_n(15) %>% 
  arrange(desc(n))
```

## Exploratory Analysis
### By Item
A plot of which items sell the most.

```{r echo = FALSE}
ggplot(top_item, aes(x = reorder(Item, -n), y = n)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Seems like coffee and bread make up the most items sold

### By day of the week
A plot by day of the week shows Saturday sells the most

```{r echo = FALSE}
top_item_wday <- df %>% 
  select(weekday, Item) %>% 
  filter(Item != "NONE") %>% 
  group_by(weekday) %>% 
  tally()

ggplot(top_item_wday, aes(x = reorder(weekday, -n), y = n)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### By day of the week and n item sold
Let's see if there are any differences in which items are sold based on day of the week for each day the top 6 items are selected. 

```{r message = FALSE, echo = FALSE}
top_item_wday3 <- df %>% 
  select(weekday, Item) %>% 
  filter(Item != "NONE") %>% 
  group_by(weekday, Item) %>%
  tally() %>% 
  top_n(6)

ggplot(top_item_wday3, aes(x = reorder(weekday, -n), y = n)) +
  geom_bar(stat = "identity") +
  facet_grid(~ Item) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Initially, Medialuna appears only on Sunday as a popular item but later on I discovered this was an issue with the coding which elimated values on months where Medialuna isn't in the top 6 for that month and possible be the 7th most item sold in that month.

Here is a graph of weekday items sold for 'top 6' items, with the items specifically 
mentioned in a vector where we can run the filter through.

```{r message = FALSE, echo = FALSE}
topitems <- c("Coffee", "Bread", "Tea", "Cake", "Pastry", "Sandwich")

top_item_wday4 <- df %>% 
  select(weekday, Item) %>% 
  filter(Item %in% topitems) %>% 
  group_by(weekday, Item) %>%
  tally() 

ggplot(top_item_wday4, aes(x = reorder(weekday, -n), y = n)) +
  geom_bar(stat = "identity") +
  facet_grid(~ Item) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### By month analysis of items sold
I removed April and October since those are the start and end months of data collection which likely caused data to be incomplete for those months.

```{r message = FALSE, echo = FALSE}
top_item_month <- df %>% 
  select(month, Item) %>% 
  filter(Item != "NONE", month != "Oct", month != "Apr") %>% 
  group_by(month, Item) %>%
  tally() %>% 
  top_n(5)

ggplot(top_item_month, aes(x = reorder(month, -n), y = n)) +
  geom_bar(stat = "identity") +
  facet_grid(~ Item) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

It's odd that some months have so little sold in Pastry, Sandwhich, and Medialuna
Is this observation valid?
(This is where I discovered my issue in my coding method)

Let's graph just medialuna, pastry, and sandwich by month.

```{r message = FALSE, echo = FALSE}
low_item_month <- df %>% 
  select(month, Item) %>% 
  filter(month != "Oct", month != "Apr", Item == "Medialuna" | Item == "Pastry" | Item == "Sandwich") %>% 
  group_by(month, Item) %>%
  tally() 
ggplot(low_item_month, aes(x = reorder(month, -n), y = n)) +
  geom_bar(stat = "identity") +
  facet_grid(~ Item) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Try to graph all items per month again but filter items specifically, instead of using top_n

```{r message = FALSE, echo = FALSE}
topitems2 <- c("Coffee", "Bread", "Tea", "Cake", "Pastry", "Sandwich", "Medialuna")
top_item_month2 <- df %>% 
  select(month, Item) %>% 
  filter(month != "Oct", month != "Apr", Item %in% topitems2) %>% 
  group_by(month, Item) %>%
  tally() 

ggplot(top_item_month2, aes(x = reorder(month, -n), y = n)) +
  geom_bar(stat = "identity") +
  facet_grid(~ Item) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

The method of using top_n messes up when a month has a different top_n that others don't have.
Here we accurately see the amounts of each item.

Let's look at how many items are bought per transaction 
```{r message = FALSE, echo = FALSE}
items_per_trans <- df %>%
  select(Item, Transaction) %>% 
  group_by(Transaction) %>% 
  tally() %>% 
  arrange(desc(n))
```
